Inferring the distribution and rate of Earth-like exoplanets \\ from noisy individual-planet inferences

### 2014-05-13 document/ms.tex
words about ess

\newcommand{\smooth}{{\ensuremath{\lambda}}}

@@

\paragraph{Occurence rate parameterization \& priors}
For the remainder of this \paper, we also model the occurence rate as a
two-dimensional histogram with fixed logarithmic bins in period and radius.
When we include observational uncertainties---using \eq{money}---the maximum
likelihood result is no longer analytic.
Therefore, if we want to compute the ``best-fit'' occurence rate, we can use a
standard non-linear optimization algorithm.

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=\textwidth]{figures/smooth/results.pdf}
\end{center}
\caption{%
Inferences about the occurence rate function based on the simulated catalog
\modela.
\emph{Center:} the points with error bars show the exoplanet candidates in the
simulated incomplete catalog, the contours show the survey completeness
function (\citealt{petigura}), and the grayscale shows the median posterior
occurence surface.
\emph{Top and left:} the red dashed line shows the \emph{true} distribution
that was used to generate the catalog, the points with error bars show the
results of the inverse-detection-efficiency procedure, and the histograms are
posterior samples from the marginalized occurence rate as inferred by our
method.
\figlabel{smooth-results}}
\end{figure}

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=\textwidth]{figures/simulation/results.pdf}
\end{center}
\caption{%
The same as \fig{smooth-results} for \modelb.
\dfmtodo{UPDATE THIS FIGURE TO FINAL VERSION}
\figlabel{simulation-results}}
\end{figure}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.6\textwidth]{figures/smooth/rate.pdf}
\end{center}
\caption{%
The extrapolated rate of Earth analogs \gammaearth\ as inferred by the
different techniques.
Applying the method used by \citet{petigura} gives a constraint indicated by
the vertical black line with error bars shown as dashed lines.
The histogram is the MCMC estimate of our posterior constraint on this rate
and the true value is indicated as the thick gray vertical line.
\figlabel{smooth-rate}}
\end{figure}

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.6\textwidth]{figures/simulation/rate.pdf}
\end{center}
\caption{%
The same as \fig{smooth-rate} for \modelb.
\dfmtodo{UPDATE THIS FIGURE TO FINAL VERSION}
\figlabel{simulation-rate}}
\end{figure}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=\textwidth]{figures/results/results.pdf}
\end{center}
\caption{%
The same as \fig{smooth-results} when applied to the observed data from
\citet{petigura}.
\figlabel{real-results}}
\end{figure}

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.6\textwidth]{figures/results/rate.pdf}
\end{center}
\caption{%
The same as \fig{smooth-rate} but applied to the catalog from \citet{petigura}.
\figlabel{real-rate}}
\end{figure}

@@

@@

@@

>>>>

\newcommand{\mean}{{\ensuremath{\mu}}}
\newcommand{\smooth}{{\ensuremath{\lambda}}}
\newcommand{\smoothpars}{{\ensuremath{\bvec{\smooth}}}}
\newcommand{\cov}{{\ensuremath{\mathrm{K}}}}

@@

\paragraph{Occurence rate parameterization \& priors}
For the remainder of this \paper, we also model the occurence rate as a
two-dimensional histogram with fixed logarithmic bins in period and radius.
When we include observational uncertainties---using \eq{money}---the maximum
likelihood result is no longer analytic.
Therefore, if we want to compute the ``best-fit'' occurence rate, we can use a
standard non-linear optimization algorithm.
In the regions of parameter space that we tend to care about, the completeness
is low and there are only a few observations with large uncertainties.
In this case, we're especially interested in probabilistic constraints on the
occurence rate; not just the best-fit model.
To do this, we apply a prior\footnote{This is a slight misnomer because
we also we also marginalize over the parameters of this model.} $p(\ratepars)$
on the occurence rate parameters and generate samples from the posterior
probability
\begin{eqnarray}\eqlabel{posterior}
p(\ratepars\,|\,\{\data_k\}) &\propto&
    p(\ratepars)\,p(\{\data_k\}\,|\,\ratepars)
\end{eqnarray}
using Markov chain Monte Carlo (MCMC).

Given our model of the rate function as a collection of bins, we choose to
model the prior as a Gaussian process (\citealt{gp,gibson-gp}) on the
logarithmic bin heights.
This model encodes our prior belief that, on the grid scale that we consider,
the occurence rate should be smooth but it is otherwise very flexible about
the form of the function.
Mathematically, the prior is written as
\begin{eqnarray}
p(\ratepars) &=& p(\ratepars\,|\,\mean,\,\smooth) \\
&=& \mathcal{N} \left[\ln\ratepars;\,\mean\,\bvec{I},\,
\cov(\{\binarea_j\},\,\smoothpars)\right]
\end{eqnarray}
where $\mathcal{N}(\cdot;\,\mean\,\bvec{I},\,\cov)$ is a $J$-dimensional
Gaussian with a constant mean \mean\ and covariance matrix \cov\ that depends
on the bin locations $\{\binarea_j\}$ and a set of hyperparameters
$\smoothpars = (\smooth_0,\,\smooth_\period,\,\smooth_\radius)$.
The covariance function that we use is an anisotropic, axis-aligned
exponential-squared kernel so elements of the matrix are
\begin{eqnarray}
\cov_{ij} &=& \smooth_0\,\exp\left(-\frac{1}{2}\,
    [\binarea_i-\binarea_j]^\mathrm{T}\,\Sigma^{-1}\,[\binarea_i-\binarea_j]
\right)
\end{eqnarray}
where $\Sigma^{-1}$ is the diagonal matrix
\begin{eqnarray}
\Sigma^{-1} &=& \left(\begin{array}{cc}
1/\smooth_\period & 0 \\
0 & 1/\smooth_\radius
\end{array}\right) \quad.
\end{eqnarray}

As well as being a very flexible prior, this Gaussian process model is also
appealing because it is very well studied in the machine learning literature.
In particular, there is a very efficient algorithm---elliptical slice sampling
(ESS; \citealt{ess})---for generating samples from the posterior probability
function in \eq{posterior}.
As mentioned above, we actually go one step further and assume uniform
(hyper-)priors on $\ln\mean$ and $\ln\smoothpars$ in order to sample from the
joint posterior
\begin{eqnarray}
p(\ratepars,\,\mean,\,\smoothpars\,|\,\{\data_k\})
\end{eqnarray}
using Algorithm 1 from \citet{ess-hyper}.

@@

\begin{figure}[p]
\begin{center}
\includegraphics[width=\textwidth]{figures/smooth/results.pdf}
\end{center}
\caption{%
Inferences about the occurence rate function based on the simulated catalog
\modela.
\emph{Center:} the points with error bars show the exoplanet candidates in the
simulated incomplete catalog, the contours show the survey completeness
function (\citealt{petigura}), and the grayscale shows the median posterior
occurence surface.
\emph{Top and left:} the red dashed line shows the \emph{true} distribution
that was used to generate the catalog, the points with error bars show the
results of the inverse-detection-efficiency procedure, and the histograms are
posterior samples from the marginalized occurence rate as inferred by our
method.
\figlabel{smooth-results}}
\end{figure}

\begin{figure}[p]
\begin{center}
\includegraphics[width=\textwidth]{figures/simulation/results.pdf}
\end{center}
\caption{%
The same as \fig{smooth-results} for \modelb.
\dfmtodo{UPDATE THIS FIGURE TO FINAL VERSION}
\figlabel{simulation-results}}
\end{figure}

@@

\begin{figure}[p]
\begin{center}
\includegraphics[width=0.6\textwidth]{figures/smooth/rate.pdf}
\end{center}
\caption{%
The extrapolated rate of Earth analogs \gammaearth\ as inferred by the
different techniques.
Applying the method used by \citet{petigura} gives a constraint indicated by
the vertical black line with error bars shown as dashed lines.
The histogram is the MCMC estimate of our posterior constraint on this rate
and the true value is indicated as the thick gray vertical line.
\figlabel{smooth-rate}}
\end{figure}

\begin{figure}[p]
\begin{center}
\includegraphics[width=0.6\textwidth]{figures/simulation/rate.pdf}
\end{center}
\caption{%
The same as \fig{smooth-rate} for \modelb.
\dfmtodo{UPDATE THIS FIGURE TO FINAL VERSION}
\figlabel{simulation-rate}}
\end{figure}

@@

\begin{figure}[p]
\begin{center}
\includegraphics[width=\textwidth]{figures/results/results.pdf}
\end{center}
\caption{%
The same as \fig{smooth-results} when applied to the observed data from
\citet{petigura}.
\figlabel{real-results}}
\end{figure}

\begin{figure}[p]
\begin{center}
\includegraphics[width=0.6\textwidth]{figures/results/rate.pdf}
\end{center}
\caption{%
The same as \fig{smooth-rate} but applied to the catalog from \citet{petigura}.
\figlabel{real-rate}}
\end{figure}

@@

\bibitem[Gibson \etal(2012)]{gibson-gp}
Gibson, N.~P., Aigrain, S., Roberts, S., \etal\ 2012, \mnras, 419, 2683

@@

\bibitem[Murray \etal(2010)]{ess}
Murray, I., Prescott Adams, R., \& MacKay, D.~J.~C.\ 2010, JMLR: W\&CP, 9, 541
(\arxiv{1001.0175})

\bibitem[Murray \& Prescott Adams(2010)]{ess-hyper}
Murray, I., \& Prescott Adams, R.\ 2010, Advances in Neural Information
Processing Systems, 23, 1723
(\arxiv{1006.0868})

@@

\bibitem[Rasmussen \& Williams(2006)]{gp}
Rasmussen, C.~E. \& Williams, C.~K.~I.\ 2006
Gaussian Processes for Machine Learning, MIT Press
(\href{http://www.gaussianprocess.org/gpml/}{online})
